---
title: "theme_pub vs e_theme_pub"
subtitle: "ggplot on the left, echarts4r on the right. base_size = 8. Panels are 360×360 px."
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    fig-width: 3.75
    fig-height: 3.75
    fig-dpi: 96
    fig-retina: 1
---

```{=html}
<style>
html, body { overflow: auto !important; }
/* Keep both panels at 360 CSS px so a retina ggplot is not
   downscaled while the echarts canvas keeps a fixed fontSize. */
div.columns {
  display: flex;
  flex-wrap: nowrap;
  gap: 0.5rem;
  overflow-x: auto;
}
div.column {
  flex: 0 0 360px;
  width: 360px;
  max-width: 360px;
  overflow: visible;
}
div.column img,
div.column .html-widget,
div.column .echarts4r,
div.column .html-widget-static-bound,
div.column .quarto-figure,
div.column .quarto-float {
  width: 360px !important;
  height: 360px !important;
  max-width: 360px !important;
}
</style>
```

```{r}
#| label: setup
#| include: false
pkgload::load_all(".", quiet = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(echarts4r)
library(scales)
library(ggrepel)
library(jsonlite)

base_size = 8
plot.px = 360
cap = "Optional caption giving more info, X handle, or shameless promotion of pubtheme"
sub = "Optional Subtitle In Upper Lower"
ttl = "Title in Upper Lower"
xlab = "Horizontal Axis Label in Upper Lower"
ylab = "Vertical Axis Label in Upper Lower"

# Scatter
dg.sc = mtcars |>
  select(wt, mpg, cyl) |>
  mutate(Cylinders = as.factor(cyl)) |>
  rename(MPG = mpg)

# Line (README: days + value01, all series)
dg.ln = economics_long |>
  mutate(
    name = toupper(variable),
    days = as.numeric(date - min(date)),
    value = value01
  ) |>
  select(date, days, name, value)

# Histogram
dg.hist = economics |>
  filter(date <= "2008-01-01") |>
  rename(value = unemploy)

# Bar
dg.bar = mtcars |>
  summarise(value = mean(mpg), .by = cyl) |>
  mutate(
    name = paste0(cyl, "-cylinder"),
    name = factor(name, levels = c("8-cylinder", "6-cylinder", "4-cylinder"))
  )
x.max = 30

# Grid / calendar month-day
dg.grid = airquality |>
  mutate(
    Month = month.abb[Month],
    Month = factor(Month, levels = rev(month.abb[month.abb %in% Month])),
    Day = factor(Day, levels = 1:31)
  ) |>
  rename(x = Day, y = Month, value = Temp) |>
  complete(x, y, fill = list(value = NA))

# Correlation
corr = round(cor(mtcars[, sort(colnames(mtcars))]), 2)
dg.corr = corr |>
  as.data.frame() |>
  rownames_to_column(var = "x") |>
  pivot_longer(cols = -x, names_to = "y", values_to = "value") |>
  mutate(
    x = factor(x, levels = rev(sort(unique(x)))),
    y = factor(y, levels = sort(unique(y)))
  )

# Calendar
df.cal = data.frame(
  date = seq.Date(as.Date("2024-01-01"), as.Date("2024-12-31"), by = "day"),
  value = rep(1:7, length.out = 366)
)

# Slope
dg.slope = economics_long |>
  mutate(
    name = toupper(variable),
    x = as.character(date),
    y = value01
  ) |>
  select(x, name, y) |>
  filter(x %in% range(x), name != "PCE") |>
  mutate(time = ifelse(x == min(x), "", "end")) |>
  pivot_wider(names_from = time, values_from = c(x, y), names_sep = "") |>
  mutate(
    diff = yend - y,
    label1 = paste0(name, ", ", round(y, 2)),
    label2 = paste0(round(yend, 2), " (", sprintf("%+.2f", diff), "), ", name)
  )
dg.slope.long = bind_rows(
  dg.slope |> transmute(name, x = x, y = y, label = label1),
  dg.slope |> transmute(name, x = xend, y = yend, label = label2)
)

# Lollipop
dg.pop = airquality |>
  mutate(Month = factor(month.abb[Month], levels = rev(month.abb))) |>
  summarise(value = mean(Temp), .by = Month) |>
  rename(name = Month)

# Discrete lollipop
dg.popd = data.frame(name = 0:10, value = dbinom(0:10, 10, 0.5))

# Barbell
dg.barb = airquality |>
  mutate(Month = factor(month.abb[Month], levels = rev(month.abb))) |>
  summarise(x = min(Temp), xend = max(Temp), .by = Month) |>
  rename(name = Month)

# Dot / errorbar
d.std = mtcars |>
  mutate(wt = scale(wt), cyl = scale(cyl), disp = scale(disp), hp = scale(hp))
m1 = lm(mpg ~ wt + cyl + disp + hp, data = d.std)
dg.dot = summary(m1)$coefficients |>
  as.data.frame() |>
  rownames_to_column(var = "var") |>
  rename(coef = Estimate, se = `Std. Error`) |>
  filter(var != "(Intercept)") |>
  mutate(
    var = toupper(var),
    lo = coef - se,
    hi = coef + se,
    label.se = paste0(sprintf("%.1f", coef), " \u00B1 ", sprintf("%.1f", se))
  )

# Facet
dg.facet = mtcars |>
  select(wt, mpg, cyl) |>
  mutate(name = paste0(cyl, "-cylinder"))

# Timeline
current.year = as.numeric(format(Sys.Date(), "%Y"))
dg.time = data.frame(
  date = as.Date(c(
    paste0(current.year, c("-01-01", "-07-04", "-12-25", "-12-31")),
    paste0(current.year + 1, c("-01-01", "-07-04", "-12-25", "-12-31"))
  )),
  text = rep(
    c("New Year's Day", "Independence Day", "Christmas Day", "New Year's Eve"),
    2
  )
) |>
  mutate(
    text = paste0(text, ", ", date),
    name = case_when(
      grepl("Christmas", text) ~ "Christmas",
      grepl("Indep", text) ~ "Indep",
      TRUE ~ "Other"
    ),
    name = factor(name, levels = c("Christmas", "Indep", "Other")),
    x = 0
  )
time.breaks = as.Date(paste0(current.year + 0:2, "-01-01"))

# Map
dg.map = data.frame(
  city = c("New York", "Los Angeles"),
  lat = c(40.7128, 34.0522),
  lon = c(-74.0060, -118.2437)
)

# Hex
dg.hex = economics

# Colors / facet grid
dg.col = mtcars |>
  mutate(
    gear = paste0(gear, " gear"),
    cyl = paste0(cyl, " cyl"),
    name = paste0(gear, ", ", cyl)
  )

# Horizontal category plots (lollipop, barbell, dot, errorbar) without
# e_flip_coords(), which breaks e_error_bar's custom renderer.
hcat.cats = function(x) {
  if (is.factor(x)) as.character(levels(droplevels(x))) else unique(as.character(x))
}
hcat.yval = function(y) {
  if (is.numeric(y)) as.numeric(y) else as.character(y)
}
hcat.init = function(df, xcol, ycol, xlim = NULL, xname = NULL) {
  cats = hcat.cats(df[[ycol]])
  e = echarts4r::e_charts_(df, xcol, width = plot.px, height = plot.px)
  e$x$opts$xAxis = list(type = "value")
  if (!is.null(xname)) e$x$opts$xAxis$name = xname
  if (!is.null(xlim)) {
    e$x$opts$xAxis$min = min(xlim)
    e$x$opts$xAxis$max = max(xlim)
  }
  # Category y so stems sit on the same row as the label and grid
  # line. Leading "" keeps the first level off the x-axis.
  e$x$opts$yAxis = list(
    type = "category",
    data = c("", cats),
    boundaryGap = FALSE
  )
  e
}
hcat.seg = function(e, x0, y, x1, color, lwd, symbol = "none", size = 8) {
  extra = lapply(seq_along(x0), function(i) {
    yi = hcat.yval(y[i])
    list(
      type = "line",
      data = list(
        list(as.numeric(x0[i]), yi),
        list(as.numeric(x1[i]), yi)
      ),
      symbol = symbol,
      symbolSize = size,
      showSymbol = !identical(symbol, "none"),
      lineStyle = list(color = color, width = lwd),
      itemStyle = list(color = color),
      legendHoverLink = FALSE,
      silent = TRUE,
      z = 2
    )
  })
  e$x$opts$series = c(e$x$opts$series, extra)
  e
}
hcat.pts = function(e, x, y, color, size, labels = NULL) {
  dat = lapply(seq_along(x), function(i) {
    pt = list(value = list(as.numeric(x[i]), hcat.yval(y[i])))
    if (!is.null(labels)) {
      pt$label = list(
        show = TRUE,
        position = "right",
        formatter = as.character(labels[i]),
        color = pubtextgray
      )
    }
    pt
  })
  s = list(
    type = "scatter",
    data = dat,
    symbol = "circle",
    symbolSize = size,
    itemStyle = list(color = color),
    z = 3
  )
  e$x$opts$series = c(e$x$opts$series, list(s))
  e
}
hcat.cap = function(e, x, y, half = 0.15, color, lwd) {
  dat = lapply(seq_along(x), function(i) {
    list(value = list(as.numeric(x[i]), hcat.yval(y[i])))
  })
  s = list(
    type = "scatter",
    data = dat,
    symbol = "rect",
    symbolSize = c(max(lwd, 1), 8),
    itemStyle = list(color = color),
    silent = TRUE,
    legendHoverLink = FALSE,
    z = 2
  )
  e$x$opts$series = c(e$x$opts$series, list(s))
  e
}
hcat.vline = function(e, x, cats, color = pubmediumgray, lwd = 1) {
  s = list(
    type = "line",
    data = list(
      list(as.numeric(x), cats[1]),
      list(as.numeric(x), cats[length(cats)])
    ),
    symbol = "none",
    lineStyle = list(color = color, width = lwd),
    silent = TRUE,
    legendHoverLink = FALSE,
    z = 1
  )
  e$x$opts$series = c(e$x$opts$series, list(s))
  e
}
usa.states.geo = function() {
  md = ggplot2::map_data("state")
  rings = split(md, md$group)
  features = lapply(unname(rings), function(d) {
    coords = Map(function(lon, lat) c(as.numeric(lon), as.numeric(lat)), d$long, d$lat)
    if (!identical(coords[[1]], coords[[length(coords)]])) {
      coords = c(coords, coords[1])
    }
    list(
      type = "Feature",
      properties = list(name = as.character(d$region[1])),
      geometry = list(type = "Polygon", coordinates = list(unname(coords)))
    )
  })
  list(type = "FeatureCollection", features = unname(features))
}
hex.symbol = "path://M0,-1L0.866,-0.5L0.866,0.5L0,1L-0.866,0.5L-0.866,-0.5Z"
```

Use the table of contents to jump to a plot type. Every README `theme_pub` type is here. Code is folded; expand a chunk if you want the source.

The line, bar, histogram, grid, calendar, slope, lollipop, barbell, dot, error bar, facet, timeline, map, hex, pairs, and correlation examples match the README templates as closely as echarts4r allows.

## Scatter

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "scatter")`**

```{r}
#| label: fig-gg-scatter
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.sc, aes(x = wt, y = MPG, color = Cylinders)) +
  geom_point() +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = ylab)
pub(g, type = "scatter", base_size = base_size, xlim = c(0, 6), ylim = c(0, 40))
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "scatter")`**

```{r}
#| label: fig-ec-scatter
dg.sc |>
  group_by(Cylinders) |>
  e_charts(wt, width = plot.px, height = plot.px) |>
  e_scatter(MPG) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_y_axis(name = ylab) |>
  e_theme_pub(
    type = "scatter",
    base_size = base_size,
    caption = cap,
    xlim = c(0, 6),
    ylim = c(0, 40)
  )
```

:::
::::

## Line

README line plot: `days` on x, `value01` on y, all series, `ylim = c(0, 1)`.

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "line")`**

```{r}
#| label: fig-gg-line
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.ln, aes(x = days, y = value, color = name)) +
  geom_line() +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = ylab, color = "Legend Label") +
  guides(color = guide_legend(nrow = 2))
pub(g, type = "line", base_size = base_size, ylim = c(0, 1))
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "line")`**

```{r}
#| label: fig-ec-line
dg.ln |>
  group_by(name) |>
  e_charts(days, width = plot.px, height = plot.px) |>
  e_line(value) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_y_axis(name = ylab) |>
  e_theme_pub(
    type = "line",
    base_size = base_size,
    caption = cap,
    ylim = c(0, 1)
  )
```

:::
::::

## Line with date

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "line")`**

```{r}
#| label: fig-gg-linedate
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.ln, aes(x = date, y = value, color = name)) +
  geom_line() +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = ylab, color = "Legend Label") +
  guides(color = guide_legend(nrow = 2))
pub(
  g,
  type = "line",
  base_size = base_size,
  ylim = c(0, 1),
  xbreaks = as.Date(c("1970-01-01", "1990-01-01", "2010-01-01")),
  xlabels = function(x) format(x, "%b%e, %Y")
)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "line")`**

```{r}
#| label: fig-ec-linedate
dg.ln |>
  group_by(name) |>
  e_charts(date, width = plot.px, height = plot.px) |>
  e_line(value) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_y_axis(name = ylab) |>
  e_theme_pub(
    type = "line",
    base_size = base_size,
    caption = cap,
    ylim = c(0, 1),
    xbreaks = as.Date(c("1970-01-01", "1990-01-01", "2010-01-01")),
    xlabels = "{MMM}{d}, {yyyy}"
  )
```

:::
::::

## Histogram

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "hist")`**

```{r}
#| label: fig-gg-hist
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.hist, aes(x = value)) +
  geom_histogram(fill = pubblue, color = pubbackgray, binwidth = 500) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = "Count")
pub(g, type = "hist", base_size = base_size)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "hist")`**

```{r}
#| label: fig-ec-hist
dg.hist |>
  e_charts(width = plot.px, height = plot.px) |>
  e_histogram(
    value,
    breaks = seq(
      floor(min(dg.hist$value) / 500) * 500,
      ceiling(max(dg.hist$value) / 500) * 500,
      by = 500
    ),
    legend = FALSE
  ) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_y_axis(name = "Count") |>
  e_theme_pub(type = "hist", base_size = base_size, caption = cap)
```

:::
::::

## Bar

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "bar")`**

```{r}
#| label: fig-gg-bar
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.bar, aes(x = value, y = name, label = round(value, 2))) +
  geom_col(aes(x = x.max), fill = publightgray, width = 0.8) +
  geom_col(width = 0.8) +
  geom_text(hjust = -0.1) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = NULL)
pub(g, type = "bar", base_size = base_size)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "bar")`**

```{r}
#| label: fig-ec-bar
dg.bar |>
  mutate(bg = x.max) |>
  arrange(name) |>
  e_charts(name, width = plot.px, height = plot.px) |>
  e_bar(
    bg,
    legend = FALSE,
    barWidth = "80%",
    barGap = "-100%",
    z = 1,
    itemStyle = list(color = publightgray)
  ) |>
  e_bar(
    value,
    legend = FALSE,
    barWidth = "80%",
    barGap = "-100%",
    z = 2,
    itemStyle = list(color = pubblue)
  ) |>
  e_flip_coords() |>
  e_labels(position = "right", formatter = htmlwidgets::JS(
    "function(p){ var v=p.value; var n=Array.isArray(v)?Number(v[0]):Number(v); if(!isFinite(n)) n=Array.isArray(v)?Number(v[v.length-1]):NaN; return (p.seriesIndex===1 && isFinite(n)) ? (Math.round(n*100)/100).toString() : ''; }"
  )) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_theme_pub(
    type = "bar",
    base_size = base_size,
    caption = cap,
    xlim = c(0, x.max)
  )
```

:::
::::

## Grid

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "grid")`**

```{r}
#| label: fig-gg-grid
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.grid, aes(x = x, y = y, fill = value)) +
  geom_tile(linewidth = 0.4, color = pubdarkgray) +
  scale_fill_gradient(low = pubgradgray, high = pubblue, na.value = pubmediumgray, oob = squish, breaks = c(60, 75, 90)) +
  labs(title = ttl, subtitle = sub, caption = cap, x = "Day", y = NULL, fill = "Value")
pub(g, type = "grid", base_size = base_size, xbreaks = seq(2, 32, by = 2))
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "grid")`**

```{r}
#| label: fig-ec-grid
dg.grid |>
  mutate(x = as.character(x), y = as.character(y)) |>
  e_charts(x, width = plot.px, height = plot.px) |>
  e_heatmap(y, value) |>
  e_visual_map(
    value,
    orient = "horizontal",
    inRange = list(color = c(pubgradgray, pubblue))
  ) |>
  e_title(ttl, sub) |>
  e_x_axis(name = "Day") |>
  e_theme_pub(type = "grid", base_size = base_size, caption = cap)
```

:::
::::

## Correlation

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "grid")`**

```{r}
#| label: fig-gg-corr
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.corr, aes(x = x, y = y, fill = value)) +
  geom_tile(linewidth = 0.4, color = pubdarkgray) +
  scale_fill_gradientn(
    colors = c("red4", pubred, publightred, "white", publightblue, pubblue, "navy"),
    na.value = pubmediumgray,
    oob = squish,
    breaks = c(-1, 0, 1),
    limits = c(-1, 1)
  ) +
  labs(title = ttl, subtitle = sub, caption = cap, x = "Variable", y = NULL, fill = "Value")
pub(g, type = "grid", base_size = base_size)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "grid")`**

```{r}
#| label: fig-ec-corr
dg.corr |>
  mutate(x = as.character(x), y = as.character(y)) |>
  e_charts(x, width = plot.px, height = plot.px) |>
  e_heatmap(y, value) |>
  e_visual_map(
    value,
    min = -1,
    max = 1,
    orient = "horizontal",
    inRange = list(color = c("red4", pubred, publightred, "#ffffff", publightblue, pubblue, "navy"))
  ) |>
  e_title(ttl, sub) |>
  e_x_axis(name = "Variable") |>
  e_theme_pub(type = "grid", base_size = base_size, caption = cap)
```

:::
::::

## Pairs

`ggpairs` is not an echarts chart. The right panel is one scatter with `type = "pairs"` so you can check the panel border and both grids.

:::: {.columns}
::: {.column width="50%"}
**ggplot + `theme_pub(type = "pairs")`**

```{r}
#| label: fig-gg-pairs
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
library(GGally)
dg.pairs = mtcars |>
  select(mpg, cyl, wt, carb) |>
  mutate(Cylinders = factor(cyl))
ggpairs(
  dg.pairs,
  aes(color = Cylinders, fill = Cylinders),
  columns = c("mpg", "wt", "carb"),
  diag = list(continuous = pub.density),
  title = ttl,
  xlab = xlab,
  ylab = ylab
) +
  theme_pub(type = "pairs", base_size = base_size)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "pairs")`**

```{r}
#| label: fig-ec-pairs
dg.sc |>
  group_by(Cylinders) |>
  e_charts(wt, width = plot.px, height = plot.px) |>
  e_scatter(MPG) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_y_axis(name = ylab) |>
  e_theme_pub(
    type = "pairs",
    base_size = base_size,
    caption = cap,
    xlim = c(0, 6),
    ylim = c(0, 40)
  )
```

:::
::::

## Calendar

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "cal")`**

```{r}
#| label: fig-gg-cal
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
dg.cal = df.cal |>
  complete(
    date = seq.Date(lubridate::floor_date(min(date), "month"), lubridate::ceiling_date(max(date), "month") - 1, by = "day"),
    fill = list(value = 0)
  ) |>
  mutate(
    month = factor(months(date, abbreviate = FALSE), levels = month.name),
    day.of.week = factor(
      weekdays(date, abbreviate = TRUE),
      levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")
    ),
    day = lubridate::day(date),
    year = lubridate::year(date)
  ) |>
  group_by(month, year) |>
  mutate(
    week = ifelse(day == 1 | day.of.week == "Sun", 1, 0),
    week = cumsum(week),
    day = as.character(day)
  ) |>
  complete(week, day.of.week, fill = list(value = NA, day = ""))

g = ggplot(dg.cal, aes(x = day.of.week, y = week, fill = value)) +
  geom_tile(linewidth = 0.4, color = pubdarkgray) +
  facet_wrap(~month, scales = "free_x", ncol = 4) +
  scale_fill_gradient(low = pubgradgray, high = pubblue, na.value = pubbackgray, oob = squish) +
  labs(title = ttl, subtitle = sub, caption = cap, x = NULL, y = NULL, fill = "Value")
pub(g, type = "cal", base_size = base_size, ytrans = "reverse")
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "cal")`**

```{r}
#| label: fig-ec-cal
df.cal |>
  e_charts(date, width = plot.px, height = plot.px) |>
  e_calendar(range = "2024") |>
  e_heatmap(value, coord_system = "calendar") |>
  e_visual_map(
    value,
    orient = "horizontal",
    inRange = list(color = c(pubgradgray, pubblue))
  ) |>
  e_title(ttl, sub) |>
  e_theme_pub(type = "cal", base_size = base_size, caption = cap)
```

:::
::::

## Slope

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "slope")`**

```{r}
#| label: fig-gg-slope
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.slope, aes(x = x, y = y, color = name, size = 4, linewidth = 0.25)) +
  geom_segment(aes(xend = xend, yend = yend)) +
  geom_point(show.legend = FALSE) +
  geom_point(aes(x = xend, y = yend), show.legend = FALSE) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = NULL) +
  guides(size = "none", linewidth = "none", color = "none")
pub(g, type = "slope", base_size = base_size)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "slope")`**

```{r}
#| label: fig-ec-slope
dg.slope.long |>
  group_by(name) |>
  e_charts(x, width = plot.px, height = plot.px) |>
  e_line(y, legend = FALSE) |>
  e_title(ttl, sub) |>
  e_theme_pub(type = "slope", base_size = base_size, caption = cap)
```

:::
::::

## Lollipop

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "pop")`**

```{r}
#| label: fig-gg-pop
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.pop, aes(x = value, y = name, label = round(value, 2), size = 1, linewidth = 0.75)) +
  geom_segment(aes(x = 0, xend = value, y = name, yend = name), color = pubred) +
  geom_point(color = pubred, show.legend = FALSE) +
  geom_text(hjust = -0.3) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = NULL) +
  guides(size = "none", linewidth = "none")
pub(g, type = "pop", base_size = base_size, xlim = c(0, 120))
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "pop")`**

```{r}
#| label: fig-ec-pop
dp = dg.pop |>
  mutate(name = droplevels(name))
e = hcat.init(dp, "value", "name", xlim = c(0, 120), xname = xlab)
e = hcat.seg(e, rep(0, nrow(dp)), as.character(dp$name), dp$value, pubred, 1.5)
e = hcat.pts(e, dp$value, as.character(dp$name), pubred, 8, labels = round(dp$value, 2))
e |>
  e_title(ttl, sub) |>
  e_theme_pub(type = "pop", base_size = base_size, caption = cap, xlim = c(0, 120))
```

:::
::::

## Lollipop discrete

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "pop")`**

```{r}
#| label: fig-gg-popd
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.popd, aes(x = name, y = value, label = round(value, 2), size = 5, linewidth = 0.5)) +
  geom_point(color = pubred) +
  geom_segment(aes(x = name, xend = name, y = 0, yend = value), color = pubred) +
  geom_text(vjust = -1) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = NULL) +
  guides(size = "none", linewidth = "none")
pub(g, type = "pop", base_size = base_size, xlim = c(0, 10), ylim = c(0, 0.3), xbreaks = 0:10)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "pop")`**

```{r}
#| label: fig-ec-popd
dg.popd |>
  mutate(name = as.character(name)) |>
  e_charts(name, width = plot.px, height = plot.px) |>
  e_bar(value, legend = FALSE, itemStyle = list(color = pubred)) |>
  e_scatter(value, legend = FALSE, color = pubred) |>
  e_labels(position = "top", formatter = htmlwidgets::JS(
    "function(p){ if(p.seriesType!=='scatter') return ''; var v=p.value; var n=Array.isArray(v)?Number(v[v.length-1]):Number(v); return isFinite(n)?n.toFixed(2):''; }"
  )) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_theme_pub(
    type = "pop",
    base_size = base_size,
    caption = cap,
    ylim = c(0, 0.3)
  )
```

:::
::::

## Barbell

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "pop")`**

```{r}
#| label: fig-gg-barb
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.barb, aes(x = x, y = name, size = 4, linewidth = 0.5)) +
  geom_point(color = pubred) +
  geom_point(aes(x = xend), color = pubred) +
  geom_segment(aes(xend = xend, yend = name), color = pubred) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = NULL) +
  guides(size = "none", linewidth = "none")
pub(g, type = "pop", base_size = base_size, xlim = c(0, 120))
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "pop")`**

```{r}
#| label: fig-ec-barb
db = dg.barb |>
  mutate(name = droplevels(name))
e = hcat.init(db, "x", "name", xlim = c(0, 120), xname = xlab)
e = hcat.seg(e, db$x, as.character(db$name), db$xend, pubred, 1.5, symbol = "circle", size = 8)
e |>
  e_title(ttl, sub) |>
  e_theme_pub(type = "pop", base_size = base_size, caption = cap, xlim = c(0, 120))
```

:::
::::

## Dot and whiskers

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "dot")`**

```{r}
#| label: fig-gg-dot
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.dot, aes(x = coef, y = var, size = 4, linewidth = 0.5)) +
  geom_segment(aes(x = lo, xend = hi, y = var, yend = var), color = pubred) +
  geom_point(color = pubred) +
  geom_vline(xintercept = 0, color = pubmediumgray) +
  labs(title = ttl, subtitle = sub, caption = cap, x = "Coefficient", y = NULL) +
  guides(size = "none", linewidth = "none")
pub(g, type = "dot", base_size = base_size, xlim = c(-5, 5))
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "dot")`**

```{r}
#| label: fig-ec-dot
dd = dg.dot |>
  mutate(var = factor(var, levels = sort(unique(var))))
e = hcat.init(dd, "coef", "var", xlim = c(-5, 5), xname = "Coefficient")
e = hcat.seg(e, dd$lo, as.character(dd$var), dd$hi, pubred, 1.5)
e = hcat.pts(e, dd$coef, as.character(dd$var), pubred, 8)
e = hcat.vline(e, 0, levels(dd$var))
e |>
  e_title(ttl, sub) |>
  e_theme_pub(type = "dot", base_size = base_size, caption = cap, xlim = c(-5, 5))
```

:::
::::

## Error bars

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "dot")`**

```{r}
#| label: fig-gg-err
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.dot, aes(x = coef, y = var, size = 4, linewidth = 0.5)) +
  geom_errorbar(aes(xmin = lo, xmax = hi), color = pubred, orientation = "y", width = 0.2) +
  geom_point(color = pubred) +
  geom_vline(xintercept = 0, color = pubmediumgray) +
  labs(title = ttl, subtitle = sub, caption = cap, x = "Coefficient", y = NULL) +
  guides(size = "none", linewidth = "none")
pub(g, type = "dot", base_size = base_size, xlim = c(-5, 5))
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "dot")`**

```{r}
#| label: fig-ec-err
de = dg.dot |>
  mutate(var = factor(var, levels = sort(unique(var))))
e = hcat.init(de, "coef", "var", xlim = c(-5, 5), xname = "Coefficient")
e = hcat.seg(e, de$lo, as.character(de$var), de$hi, pubred, 1.5)
e = hcat.cap(e, de$lo, as.character(de$var), half = 0.15, color = pubred, lwd = 1.5)
e = hcat.cap(e, de$hi, as.character(de$var), half = 0.15, color = pubred, lwd = 1.5)
e = hcat.pts(e, de$coef, as.character(de$var), pubred, 8)
e = hcat.vline(e, 0, levels(de$var))
e |>
  e_title(ttl, sub) |>
  e_theme_pub(type = "dot", base_size = base_size, caption = cap, xlim = c(-5, 5))
```

:::
::::

## Facet

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(facet = TRUE)`**

```{r}
#| label: fig-gg-facet
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.facet, aes(x = wt, y = mpg, color = name)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~name, nrow = 1) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = ylab)
pub(g, base_size = base_size, xlim = c(0, 6), ylim = c(0, 40), facet = TRUE)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(facet = TRUE)`**

```{r}
#| label: fig-ec-facet
dg.facet |>
  group_by(name) |>
  e_charts(wt, width = plot.px, height = plot.px) |>
  e_scatter(mpg, legend = FALSE) |>
  e_facet(
    rows = 1,
    cols = 3,
    legend_space = 0,
    margin_trbl = c(t = 2, r = 2, b = 8, l = 10),
    h_panel_space = 4
  ) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_y_axis(name = ylab) |>
  e_theme_pub(
    type = "scatter",
    facet = TRUE,
    base_size = base_size,
    caption = cap,
    xlim = c(0, 6),
    ylim = c(0, 40)
  )
```

:::
::::

## Timeline

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "timeline")`**

```{r}
#| label: fig-gg-time
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.time, aes(x = 0, y = date, color = name, label = text)) +
  geom_segment(
    aes(x = 0, xend = 0, y = min(date), yend = max(date)),
    show.legend = FALSE,
    color = publightgray
  ) +
  geom_point(show.legend = FALSE) +
  geom_label_repel(nudge_x = 1, hjust = 0, direction = "y", show.legend = FALSE) +
  labs(title = ttl, subtitle = sub, caption = cap, x = "", y = "")
pub(
  g,
  type = "timeline",
  base_size = base_size,
  xlim = c(0, 5),
  ybreaks = time.breaks,
  ytrans = "reverse2"
)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "timeline")`**

```{r}
#| label: fig-ec-time
dt = dg.time |>
  mutate(x = 0.3, y = as.numeric(date))
e = dt |>
  group_by(name) |>
  e_charts(x, width = plot.px, height = plot.px) |>
  e_scatter(y, bind = text, legend = FALSE)
e$x$opts$series = c(
  list(list(
    type = "line",
    data = list(c(0, min(dt$y)), c(0, max(dt$y))),
    symbol = "none",
    lineStyle = list(color = publightgray, width = 2),
    silent = TRUE,
    z = 1
  )),
  e$x$opts$series
)
for (i in seq_along(e$x$opts$series)) {
  if (!identical(e$x$opts$series[[i]]$type, "scatter")) next
  e$x$opts$series[[i]]$label = list(
    show = TRUE,
    position = "right",
    formatter = htmlwidgets::JS("function(p){ return p.name || ''; }")
  )
}
e$x$opts$yAxis = list(
  type = "value",
  inverse = TRUE,
  min = min(dt$y),
  max = max(dt$y),
  axisLabel = list(
    formatter = htmlwidgets::JS(
      "function(v){ var d=new Date(v*86400*1000); var m=['Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec']; return m[d.getUTCMonth()]+' '+d.getUTCDate()+', '+d.getUTCFullYear(); }"
    )
  )
)
e |>
  e_title(ttl, sub) |>
  e_theme_pub(
    type = "timeline",
    base_size = base_size,
    caption = cap,
    xlim = c(0, 5),
    ybreaks = as.numeric(time.breaks)
  )
```

:::
::::

## Map

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(type = "map")`**

```{r}
#| label: fig-gg-map
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.map, aes(x = lon, y = lat)) +
  annotation_borders("state", colour = publightgray) +
  geom_point(color = pubred) +
  labs(title = ttl, subtitle = sub, caption = cap)
pub(g, type = "map", base_size = base_size)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "map")`**

```{r}
#| label: fig-ec-map
usa.json = jsonlite::toJSON(usa.states.geo(), auto_unbox = TRUE)
dg.map |>
  e_charts(lon, width = plot.px, height = plot.px) |>
  e_map_register("usa_states", usa.json) |>
  e_geo(
    map = "usa_states",
    roam = FALSE,
    boundingCoords = list(c(-127, 23.5), c(-65.5, 50.5)),
    itemStyle = list(areaColor = pubbackgray, borderColor = publightgray),
    emphasis = list(itemStyle = list(areaColor = pubbackgray))
  ) |>
  e_scatter(lat, coord_system = "geo", legend = FALSE, color = pubred) |>
  e_title(ttl, sub) |>
  e_theme_pub(type = "map", base_size = base_size, caption = cap)
```

:::
::::

## Hex bins

`pub()` uses the default scatter type for hex bins.

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub()` + `geom_hex`**

```{r}
#| label: fig-gg-hex
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g.hex = ggplot(dg.hex, aes(x = psavert, y = uempmed)) +
  geom_hex(color = pubbackgray, bins = 20) +
  scale_fill_gradient(low = pubbackgray, high = pubblue, na.value = pubmediumgray, oob = squish, limits = c(0, 20), breaks = c(0, 10, 20)) +
  labs(title = ttl, subtitle = sub, caption = cap, x = "Personal Savings Rate", y = "Median Duration of Unemployment", fill = "Observations")
g.hex = pub(g.hex, base_size = base_size, xlim = c(0, 20), ylim = c(0, 30))
dg.hex.bins = ggplot_build(g.hex)$data[[1]] |>
  select(x, y, count, width, height)
g.hex
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(type = "scatter")`**

```{r}
#| label: fig-ec-hex
e = dg.hex.bins |>
  e_charts(x, width = plot.px, height = plot.px) |>
  e_scatter(y, count, legend = FALSE, symbol = hex.symbol) |>
  e_visual_map(
    count,
    min = 0,
    max = 20,
    orient = "horizontal",
    inRange = list(color = c(pubbackgray, pubblue))
  ) |>
  e_title(ttl, sub) |>
  e_x_axis(name = "Personal Savings Rate") |>
  e_y_axis(name = "Median Duration of Unemployment") |>
  e_theme_pub(
    type = "scatter",
    base_size = base_size,
    caption = cap,
    xlim = c(0, 20),
    ylim = c(0, 30)
  )
gr = e$x$opts$grid[[1]]
hex.w = unique(dg.hex.bins$width)[1]
hex.h = unique(dg.hex.bins$height)[1]
sx = hex.w / 20 * (plot.px - gr$left - gr$right)
sy = hex.h / 30 * (plot.px - gr$top - gr$bottom)
# Path bbox width 2; height 2*sy/sx. Uniform scale to sx then
# matches ggplot's data-unit hex size on this panel.
a = sy / sx
hex.path = sprintf(
  "path://M0,%.4fL1,%.4fL1,%.4fL0,%.4fL-1,%.4fL-1,%.4fZ",
  a, a / 2, -a / 2, -a, -a / 2, a / 2
)
e$x$opts$series = lapply(e$x$opts$series, function(s) {
  if (is.list(s) && identical(s$type, "scatter")) {
    s$symbol = hex.path
    s$symbolSize = sx
    if (is.null(s$itemStyle)) s$itemStyle = list()
    s$itemStyle$borderColor = pubbackgray
    s$itemStyle$borderWidth = 0.5
  }
  s
})
vm = e$x$opts$visualMap
if (!is.null(vm$min) || !is.null(vm$type)) {
  e$x$opts$visualMap$min = 0
  e$x$opts$visualMap$max = 20
} else if (length(vm) > 0 && is.list(vm[[1]])) {
  e$x$opts$visualMap[[1]]$min = 0
  e$x$opts$visualMap[[1]]$max = 20
}
e
```

:::
::::

## Colors (`cb.pal`)

:::: {.columns}
::: {.column width="50%"}
**ggplot + `pub(facet = TRUE)` + `cb.pal`**

```{r}
#| label: fig-gg-colors
#| fig-width: 3.75
#| fig-height: 3.75
#| out-width: "360px"
#| out-height: "360px"
g = ggplot(dg.col, aes(x = wt, y = mpg, color = name)) +
  geom_point() +
  facet_grid(cyl ~ gear) +
  labs(title = ttl, subtitle = sub, caption = cap, x = xlab, y = ylab, color = "Gears and Cylinders") +
  scale_color_manual(values = cb.pal) +
  guides(color = guide_legend(nrow = 3))
pub(g, base_size = base_size, xlim = c(0, 6), ylim = c(0, 40), facet = TRUE)
```

:::
::: {.column width="50%"}
**echarts4r + `e_theme_pub(colors = "cb14")`**

```{r}
#| label: fig-ec-colors
dg.col |>
  group_by(cyl, gear) |>
  e_charts(wt, width = plot.px, height = plot.px) |>
  e_scatter(mpg) |>
  e_facet(
    rows = 3,
    cols = 3,
    legend_space = 0,
    margin_trbl = c(t = 2, r = 2, b = 8, l = 10),
    h_panel_space = 4,
    v_panel_space = 5
  ) |>
  e_title(ttl, sub) |>
  e_x_axis(name = xlab) |>
  e_y_axis(name = ylab) |>
  e_theme_pub(
    type = "scatter",
    facet = TRUE,
    colors = "cb14",
    base_size = base_size,
    caption = cap,
    xlim = c(0, 6),
    ylim = c(0, 40)
  )
```

:::
::::

Images-as-labels from the README are ggplot/`ggtext` only, so they are not repeated here.
