---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.height=7, 
  fig.width=7,
  dpi=84,
  #out.width="80%",
  fig.align='center',
  fig.show='hold'
)
library(knitr)
```

# pubtheme

<!-- badges: start -->
<!-- badges: end -->

The package `pubtheme` contains a `ggplot` theme `theme_pub` for creating data journalism-style data visualizations with color palettes and formatting similar to those used by media organizations like BBC, NY Times, and ESPN. Several templates for scatter plot, line plots, etc., are provided below for easy copying/pasting. 

Organization-specific color palettes and logos can be used as well via the package `orgthemes`. See https://github.com/bmacGTPM/orgthemes.

## Installation

If you don't have the package `devtools`, install it using `install.packages('devtools')`. If you have `devtools`, you can install the GitHub version of `pubtheme` with:

``` r
devtools::install_github("bmacGTPM/pubtheme")
```

If you get an error about download method, try changing this option before installing. 

```r
options(download.file.method = 'libcurl')
```

Load the package using 

```{r, message=F}
library(pubtheme)
```

as usual. The theme will change some of your ggplot defaults the first time you use it. To change them back, restart R, or use 

```{r, eval=F}
restore.ggplot.defaults()
```

at any time. It is recommended that you update your version of `tidyverse` and especially `ggplot2` to use this package. 

## Scatter plot
```{r fig.height=7, fig.width=7}
dg = mtcars %>% 
  select(wt, mpg, cyl)

title = "Title in Upper Lower" 
g = ggplot(dg, aes(x=wt, y=mpg, color=as.factor(cyl)))+
  geom_point(aes(size=mpg))+
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower',
       y = 'Vertical Axis Label in Upper Lower')+
  scale_x_continuous(limits=c(0, 6), breaks=c(0, 3, 6), oob=squish, labels=comma)+
  scale_y_continuous(limits=c(0,40), breaks=c(0,20,40), oob=squish, labels=comma)+
  coord_cartesian(clip='off', expand=FALSE)+
  scale_size(range=c(2,6))+
  theme_pub(type='scatter', base_size = 36/3) 
print(g)

## Save to a file
gg = g +
  scale_size(range=c(6,18)) +
  theme_pub(type='scatter', base_size=36)
  

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), 
      plot=gg,
      width=20,   
      height=20,  
      units='in', 
      dpi=72)     
```

You must have a subfolder called `img` in order for the `ggsave` chunk above to work. 

Note that we used `base_size=12` when viewing the image in RStudio. Use `ggsave` and `base_size=36` when saving an image instead of exporting from the RStudio viewer. Do not adjust the `width=20` in ggsave. 

Do not change `width=20`, `units='in'`, or `dpi=72`. Height can be adjusted if desired. A square image is often preferred, so when in doubt, keep height at 20.

Upper Lower means First letter of each word is capitalized. The option `expand=FALSE` removes the default padding. The option `breaks=c(0,3,6)` sets 3 lines at left, middle, and right. You can certainly add lines if there is a reason to, but when in doubt you can stick with just 3 lines (left/middle/right) only. Similarly, for the y-axis, top/middle/bottom only.  

You'll notice a `scale_size(range=c(6,18)` when preparing the plot to be saved. 
If you are using `size` inside `aes()`, you'll need that change the scale, otherwise the points will be too small. 

## Pairs plot

```{r fig.height=7, fig.width=7}
library(GGally) ## Needed for ggpairs function
dg = mtcars %>%
  select(mpg, cyl, wt, carb) %>%
  mutate(cyl=factor(cyl))

title = 'Title in Upper Lower'
g = ggpairs(dg, 
            columns = c('mpg', 'wt', 'carb'),
            aes(color = cyl, 
                fill  = cyl), 
            diag = list(continuous = pub.density)) +
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower',
       y = 'Vertical Axis Label in Upper Lower')+
  coord_cartesian(clip='off', expand=FALSE)+
  theme_pub(type='pairs', base_size=36/3)
print(g)

## Save to a file
gg = g +
  theme_pub(type='pairs', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), 
      plot=gg,
      width=20,   
      height=20,  
      units='in', 
      dpi=72)  
```


## Line plot
Since `name` and `value` will be more common column names when using these in the wild, we'll rename some columns to be name and value, to simplify copying and pasting later. 

Continuous variables for `x` seem to be more common, so we'll convert date to days for this example, so that copying and pasting later becomes easier. We'll use `toupper()` to avoid using variables with all lowercase letters in the legend.

```{r}
dg = economics_long %>%
  mutate(name = toupper(variable), 
         days = as.numeric(date - min(date)), 
         value= value01) %>%
  select(days, name, value)
head(dg)

title = "Title in Upper Lower" 
g = ggplot(dg, aes(x=days, y=value, color=name))+
  geom_line()+
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower', 
       y = 'Vertical Axis Label in Upper Lower')+  
  scale_x_continuous(labels=comma) + 
  scale_y_continuous(labels=comma, limits=c(0,1), breaks=c(0, .5, 1))+ 
  coord_cartesian(clip='off', expand=FALSE)+
  theme_pub(type='line', base_size=36/3) 
print(g)

## Save to a file
gg = g + 
  theme_pub(type='line', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), ## must have a subfolder called 'img'
       plot=gg,
       width=20,   
       height=20,  
       units='in', 
       dpi=72)     
```

Note that once again we set breaks for the y-axis at the top, middle, and bottom. 

## Histogram
```{r}
dg = economics %>%
  filter(date<='2008-01-01') %>%
  rename(value=unemploy)

title = "Title in Upper Lower" 
g  = ggplot(dg, aes(x=value))+
  geom_histogram(fill=pubred, color=pubbackgray, binwidth=500) + ## set a reasonable binwidth
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower', ## Required.
       y = 'Count')+  ## Often don't need to change.
  scale_x_continuous(labels=comma)+
  scale_y_continuous(labels=comma, expand = c(0,0))+
  theme_pub(type='hist', base_size=36/3) 
print(g)

gg = g + 
  theme_pub(type='hist', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), 
       plot = gg,
       width=20,   ## do not change
       height=20,  
       units='in', ## do not change
       dpi=72)     ## do not change
```

Here `binwidth` will almost surely need to be changed for your data. 

## Bar plot

We'll use the `mtcars` data again, with some modifications. For cylinder, we'll add
`-cylinder` to the numbers so it looks nice. We'll then convert it a factor so that we can specify the order. We'll also create a column called `max` which controls the length of the lightgray bars in the background. Finally, we'll rename `cyl` and `mpg` to `name` and `value`, which will be convenient when copying and pasting this code. 


```{r fig.height=7*.75}
dg = mtcars %>%
  group_by(cyl) %>%
  summarise(mpg = mean(mpg)) %>%
  mutate(cyl = paste0(cyl, '-cylinder'),
         cyl = factor(cyl, levels=c('8-cylinder', '6-cylinder', '4-cylinder')), 
         max = 35) %>% ## for background cars
  rename(name  = cyl, 
         value = mpg) 

title = "Title in Upper Lower" 
g = ggplot(dg, aes(x=value, y=name))+
  geom_bar(stat='identity', aes(x=max), color=NA, fill=publightgray, width=0.8)+ ## optional background bars. 
  geom_bar(stat='identity', fill=pubred, color=NA, width=0.8)+ 
  geom_text(aes(label=round(value,2)), hjust=-0.1)+ ## optional numbers with reasonable number of digits
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower', ## Optional. 
       y = NULL)+  ## Optional. Upper Lower.
  scale_x_continuous(limits=c(0,35))+ 
  theme_pub(type='bar', base_size=36/3) 
print(g)

gg = g + 
  theme_pub(type='bar', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), ## must have a subfolder called 'img'
       plot = gg,
       width=20,   ## do not change
       height=15,  
       units='in', ## do not change
       dpi=72)     ## do not change

```

If you are using digits next to the bars, you can increase the `max` so the text fits. 

You may want to remove `x` and `y` axis titles.  If not used, use `x=NULL` and/or `y=NULL` above. Do not use `x=''` and/or `y=''`. If you do use axes titles, they should be in Upper Lower. 


## Grid plot
```{r fig.height=3.5}
dg = airquality %>%
  mutate(Month= month.abb[Month],
         Month = factor(Month, levels=rev(month.abb))) %>% 
  rename(name  = Day,
         name2 = Month,
         value = Temp)

title = "Title in Upper Lower" 
g = ggplot(dg, aes(x=name, y=name2, fill=value))+
  geom_tile(size=0.4, show.legend = F) +
  scale_fill_gradient(low = pubbackgray,
                      high = pubred,
                      na.value = 'white',
                      oob=squish) +
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Day (Optional Axis Label in Upper Lower)', 
       y = NULL)+  ## Optional. 
  geom_vline(xintercept=1:(length(unique(dg$name ))+1)-.5, color=pubdarkgray, linewidth=0.2)+ # vert  lines between each square
  geom_hline(yintercept=1:(length(unique(dg$name2))+1)-.5, color=pubdarkgray, linewidth=0.2)+ # horiz lines 
  scale_x_continuous(expand = c(0, 0), position='top', breaks=seq(2,30,by=2))+
  scale_y_discrete(  expand = c(0, 0)) +
  theme_pub(type='grid', base_size=36/3) 

print(g)

gg = g + 
  theme_pub(type='grid', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), 
       plot=gg,
       width=20,   ## do not change
       height=10,  
       units='in', ## do not change
       dpi=72)     ## do not change
```


## Lollipop plot
We'll make a horizontal lollipop plot. It can be used an alternative to a bar plot, and it is often preferred, especially when the visualization will eventually be printed since the lollipop plot used far less ink/toner. 

```{r fig.height=4}
dg = airquality %>%
  mutate(Month= month.abb[Month],
         Month = factor(Month, levels=rev(month.abb))) %>% 
  group_by(Month) %>%
  summarise(Temp=mean(Temp)) %>%
  rename(name  = Month,
         value = Temp)


title = "Title in Upper Lower" 
g = ggplot(dg, aes(x=value, y=name))+
  geom_point(color=pubred)+ 
  geom_segment(aes(x=0, xend=value, y=name, yend=name), color=pubred)+
  geom_text(aes(label=round(value,2)), hjust=-0.3)+ ## optional numbers with reasonable number of digits
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower', ## Optional. 
       y = NULL)+  ## Optional. Upper Lower.
  scale_x_continuous(limits=c(0,120), expand=c(0,0))+
  coord_cartesian(clip='off', expand=FALSE)+
  theme_pub(type='pop', base_size=36/3) 
print(g)

gg = g + 
  theme_pub(type='pop', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), ## must have a subfolder called 'img'
       plot = gg,
       width=20,   ## do not change
       height=15,  
       units='in', ## do not change
       dpi=72) 
```

## Lollipop for discrete distributions
This still uses `type='pop'` but we'll use a different template for copying/pasting.

```{r}
dg = data.frame(name=0:10, value=dbinom(0:10, 10, .5))

title = "Title in Upper Lower" 
g = ggplot(dg, aes(x=name, y=value))+
  geom_point(color=pubred)+ 
  geom_segment(aes(x=name, xend=name, y=0, yend=value), color=pubred)+
  geom_text(aes(label=round(value,2)), vjust=-0.5)+ ## optional numbers with reasonable number of digits
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower', ## Optional. 
       y = NULL)+  ## Optional. Upper Lower.
  scale_x_continuous(limits=c(0,10), expand=c(0,0), breaks=0:10)+
  scale_y_continuous(limits=c(0,.3), breaks=c(0, .1, .2, .3))+
  coord_cartesian(clip='off', expand=FALSE)+
  theme_pub(type='pop', base_size=36/3) 
print(g)

gg = g + 
  theme_pub(type='pop', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), ## must have a subfolder called 'img'
       plot = gg,
       width=20,   ## do not change
       height=15,  
       units='in', ## do not change
       dpi=72) 

```

## Barbell plot
This still uses `type='pop'` but we'll use a different template for copying/pasting.

```{r fig.height=4}
dg = airquality %>%
  mutate(Month= month.abb[Month],
         Month = factor(Month, levels=rev(month.abb))) %>% 
  group_by(Month) %>%
  summarise(x   =min(Temp), 
            xend=max(Temp)) %>%
  rename(name  = Month)


title = "Title in Upper Lower" 
g = ggplot(dg, aes(y=name))+
  geom_point(aes(x=x   ), color=pubred)+ 
  geom_point(aes(x=xend), color=pubred)+ 
  geom_segment(aes(x=x, xend=xend, yend=name), color=pubred)+
  geom_text(aes(x=x,    label=round(x   ,2)), hjust=1.3)+ ## optional numbers with reasonable number of digits
  geom_text(aes(x=xend, label=round(xend,2)), hjust=-0.3)+ ## optional numbers with reasonable number of digits
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower', ## Optional. 
       y = NULL)+  ## Optional. Upper Lower.
  scale_x_continuous(limits=c(0,120), expand=c(0,0))+
  coord_cartesian(clip='off', expand=FALSE)+
  theme_pub(type='pop', base_size=36/3) 
print(g)

gg = g + 
  theme_pub(type='pop', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), ## must have a subfolder called 'img'
       plot = gg,
       width=20,   ## do not change
       height=15,  
       units='in', ## do not change
       dpi=72) 
```




## Dot and Whiskers Plot

It is often desirable to visualize regression coefficients using a dot and whiskers plot instead of showing a table of coefficients and standard errors.  We'll make up a regression model and demonstrate that here. 

This still uses `type='pop'` but we'll use a different template for copying/pasting.

```{r fig.height=7*10/20}
## standardize predictors so they are roughly the same scale
d = mtcars %>%
  mutate(wt = scale(wt),  
         cyl=scale(cyl), 
         disp=scale(disp), 
         hp = scale(hp))

m1 = lm(mpg ~  wt + cyl + disp + hp, data=d)
#summary(m1)

dg = summary(m1)$coefficients %>% 
  as.data.frame() %>%
  rownames_to_column(var = 'var') %>%
  rename(coef=Estimate, se=`Std. Error`) %>%
  select(var, coef, se) %>%
  mutate(var = toupper(gsub('[(]|[)]', '', var))) %>%
  filter(var!='INTERCEPT')

title = "Title in Upper Lower" 
g = ggplot(dg, aes(x=coef, y=var))+
  geom_segment(aes(x    = coef-se, 
                   xend = coef+se,
                   y    = var, 
                   yend = var), color=pubred)+
  geom_point(color=pubred) +
  geom_vline(xintercept=0, color=pubmediumgray)+
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Coefficient', 
       y = NULL)+  ## Optional. 
  scale_x_continuous(limits=c(-5,5))+
  scale_y_discrete(expand=c(0,0)) +
  coord_cartesian(clip='off')+
  theme_pub(type='pop', base_size=36/3) 

print(g)

gg = g + 
  theme_pub(type='pop', base_size=36)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), 
       plot=gg,
       width=20,   ## do not change
       height=10,  
       units='in', ## do not change
       dpi=72)     ## do not change


```


## Faceting
We'll use our scatter plot example, but with `facet_wrap` to make separate plots for each `cyl`.

```{r fig.height=7*13/20}
dg = mtcars %>%
  select(wt, mpg, cyl) %>%
  mutate(cyl = paste0(cyl, '-cylinder')) %>%
  rename(name = cyl)

title = "Title in Upper Lower" ## to be used by ggplot and ggsave
g = ggplot(dg, aes(x=wt, y=mpg))+
  geom_point(aes(size=mpg, color=name), show.legend=F)+
  facet_wrap(~name, nrow=1) +
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = 'Horizontal Axis Label in Upper Lower',
       y = 'Vertical Axis Label in Upper Lower')+
  scale_x_continuous(limits=c(0, 6), breaks=c(0, 3, 6), oob=squish, labels=comma_format(accuracy = 1))+
  scale_y_continuous(limits=c(0,40), breaks=c(0,20,40), oob=squish, labels=comma)+
  coord_cartesian(clip='off', expand=FALSE)+
  theme_pub(type='scatter', base_size=36/3, facet=T)
print(g)

gf = g # save this 'g' with 'f'aceting for later

gg = g +
  theme_pub(type='scatter', base_size=36, facet=T)

ggsave(filename=paste0("img/", gsub("%", " Perc", title), ".jpg"), ## must have a subfolder called 'img'
       plot=gg, ## change range=c(6,18) when base_size=36
       width=20,   ## do not change
       height=13,  ## can change if desired. Here, 14 was chosen so that each subplot is square
       units='in', ## do not change
       dpi=72)     ## do not change

```

## Timeline
Let's make up some data. Let's also define a new "reverse" function.
```{r  fig.height=7, fig.width=5}
current.year = as.numeric(format(Sys.Date(), "%Y"))
dg = data.frame(date=as.Date(c(paste0(current.year, 
                                   c('-01-01', 
                                     '-07-04', 
                                     '-12-25', 
                                     '-12-31')),
                             paste0(current.year+1, 
                                    c('-01-01', 
                                      '-07-04', 
                                      '-12-25', 
                                      '-12-31')))), 
                text=c("New Year's Day", 
                       "Independence Day",
                       "Christmas Day",
                       "New Year's Eve", 
                       "New Year's Day", 
                       "Independence Day",
                       "Christmas Day",
                       "New Year's Eve")) %>%
  mutate(text = paste0(text, ', ', date), 
         name = case_when(grepl('Christmas', text) ~ 'Christmas', 
                          grepl('Indep'    , text) ~ 'Indep', 
                          TRUE ~ 'Other'), 
         name = factor(name, levels=c('Christmas', 'Indep', 'Other')))
dg

library(ggrepel)

## Copied from https://github.com/tidyverse/ggplot2/issues/4014
reverse2_trans <- function() {
  trans_new(
    "reverse2",
    function(x) -1 * as.numeric(x), # Force values to be numeric for Date objects
    function(x) -1 * as.numeric(x)
  )
}

## Edit breaks
breaks = as.Date(c('2022-01-01', '2023-01-01', '2024-01-01'))
title = "Title in Upper Lower" ## to be used by ggplot and ggsave
g = ggplot(dg, aes(x=0, y=date, color=name))+
  geom_segment(aes(x=0, xend=0, y=min(date), yend=max(date)), 
               show.legend = F, 
               color=publightgray)+
  geom_point(show.legend=F)+
  geom_label_repel(aes(label=text),
                   nudge_x = 1,
                   hjust = 0, 
                   direction = 'y',
                   show.legend = F)+
  labs(title    = title,
       subtitle = 'Optional Subtitle In Upper Lower',
       caption  = "Optional caption, giving additional info or Twitter handle",
       x = '',
       y = '')+
  scale_x_continuous(limits=c(0, 5), expand=c(0,0))+
  scale_y_continuous(trans=c('reverse2'), 
                     breaks=as.numeric(breaks), 
                     labels=format(breaks, '%b %d, %Y'), 
                     expand=c(.1, .1))+ ## expand in case more room is needed
  coord_cartesian(clip='off')+
  theme_pub(type='timeline', base_size=12)
  
g

```


## Default colors

This theme changes your default ggplot colors to those found in `colors.r`. 
The palette for discrete color scales `scale_colour_discrete` and 
`scale_fill_discrete` consists of red, blue, gray, light red, and light blue, 
as seen in the Line Plot above. Recall that if you want to undo the changes made by this theme, you can use `restore.ggplot.defaults()` at any time.

If more than 5-6 colors are needed, a 14-color colorblind friendly palette `cb.pal` 
can be used by adding 
`+ scale_color_manual(values=cb.pal)` or 
`+ scale_fill_manual( values=cb.pal)` to a plot. For example,

```{r fig.height=7*13/20}
gf + scale_color_manual(values=cb.pal)
```

